### 2021-04-13 introduction.tex
Compacted references.

We address this research question by following the line of work from CP to Priority CP, and present
Priority GV (PGV), a~variant of GV which decouples channel creation from thread spawning, thus allowing cyclic-structured processes and restores deadlock freedom via priorities. This closes the circle of the connection between CP and GV~\cite{wadler14}, and their priority-based versions, PCP~\cite{dardhagay18} and PGV.%
In this paper we make the following main contributions:
\begin{enumerate}
	\item \textbf{Priority GV}. We present Priority GV (PGV), a session-typed functional language with priorities (\cref{sec:pgv}) and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
	PGV addresses several problems in the original GV language, most notably:
 	\begin{enumerate*}
	\item we do not require the pseudo-type $S^\sharp$; and
	\item our structural congruence is type preserving.
	\end{enumerate*}
	PGV answers our research question as it allows cyclic-structured binary session-typed processes that are deadlock free.
	\item \textbf{Updated PCP}. We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}. We remove commuting conversions and move away from reduction as cut elimination, and present a reduction relation based on structural congruence, as it is standard in process calculi.
	\item \textbf{Translation from PCP to PGV}. We present a \emph{sound and complete encoding} of PCP to PGV in~\cref{sec:pcp}, where we prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
\end{enumerate}
\end{document}

>>>>

We address this research question by following the line of work from CP to Priority CP, and present
Priority GV (PGV), a~variant of GV which decouples channel creation from thread spawning, thus allowing cyclic-structured processes and restores deadlock freedom via priorities. This closes the circle of the connection between CP and GV~\cite{wadler14}, and their priority-based versions, PCP~\cite{dardhagay18} and PGV.%
In this paper we make the following main contributions:
\begin{enumerate}
	\item \textbf{Priority GV}. We present Priority GV (PGV), a session-typed functional language with priorities (\cref{sec:pgv}) and prove subject reduction (\cref{thm:pgv-subject-reduction-confs}) and progress (\cref{thm:pgv-closed-progress-confs}).
	PGV addresses several problems in the original GV language, most notably:
 	\begin{enumerate*}
	\item we do not require the pseudo-type $S^\sharp$; and
	\item our structural congruence is type preserving.
	\end{enumerate*}
	PGV answers our research question as it allows cyclic-structured binary session-typed processes that are deadlock free.
	\item \textbf{Updated PCP}. We present an updated version of Priority CP~\cite{dardhagay18} in \cref{sec:pcp}. We remove commuting conversions and move away from reduction as cut elimination, and present a reduction relation based on structural congruence, as it is standard in process calculi.
	\item \textbf{Translation from PCP to PGV}. We present a \emph{sound and complete encoding} of PCP in PGV in~\cref{sec:pcp}, where we prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).
\end{enumerate}
\end{document}

### 2021-04-13 priority-gv.tex
Compacted references.

We present Priority GV (PGV), a session-typed functional language based on GV~\cite{wadler15,lindleymorris15} which uses priorities \`{a} la Kobayashi and Padovani~\cite{kobayashi06,padovaninovara15} to enforce deadlock freedom.
Priority GV offers a more fine-grained analysis of communication structures, and by separating channel creation form thread spawning it allows cyclic structures.
We illustrate this with two programs in PGV, \cref{ex:cycle,ex:deadlock}. Each program contains two processes---the main process, and the child process created by $\tm{\spawn}$---which communicate using \emph{two} channels. The child process receives a unit over the channel $\tm{x}/\tm{x'}$, and then sends a unit over the channel $\tm{y}/\tm{y'}$. The main process does one of two things:
\begin{enumerate*}[label=(\alph*)]
\item in \cref{ex:cycle}, it sends a unit over the channel $\tm{x}/\tm{x'}$, and then waits to receive a unit over the channel $\tm{y}/\tm{y'}$;
\item in \cref{ex:deadlock}, it does these in the opposite order, which results in a deadlock.
\end{enumerate*}
We will show that only \cref{ex:cycle} is typeable in PGV, while it is not typable in GV~\cite{wadler14} nor guaranteed to be deadlock-free by its predecessor~\cite{gayvasconcelos10}.

>>>>

We present Priority GV (PGV), a session-typed functional language based on GV~\cite{wadler15,lindleymorris15} which uses priorities \`{a} la Kobayashi and Padovani~\cite{kobayashi06,padovaninovara15} to enforce deadlock freedom.
Priority GV offers a more fine-grained analysis of communication structures, and by separating channel creation form thread spawning it allows cyclic structures.
We illustrate this with two programs in PGV, \cref{ex:cycle,ex:deadlock}. Each program contains two processes---the main process, and the child process created by $\tm{\spawn}$---which communicate using \emph{two} channels. The child process receives a unit over the channel $\tm{x}/\tm{x'}$, and then sends a unit over the channel $\tm{y}/\tm{y'}$. The main process does one of two things:
\begin{enumerate*}[label=(\alph*)]
\item in \cref{ex:cycle}, it sends a unit over the channel $\tm{x}/\tm{x'}$, and then waits to receive a unit over the channel $\tm{y}/\tm{y'}$;
\item in \cref{ex:deadlock}, it does these in the opposite order, which results in a deadlock.
\end{enumerate*}
We will show that only \cref{ex:cycle} is typeable in PGV, while it is not typable in GV~\cite{wadler14} nor guaranteed to be deadlock-free by its predecessor~\cite{gayvasconcelos12}.
